---
title: "RepDataPA2"
output: 
        html_document:
                keep_md: true
---

# Summary
This analysis is a part of the assignment in "Reprodicuble Research" course
in Coursera. 

Storms and other severe weather events can cause both public health and economic problems for communities and municipalities.
Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.
This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database.

- National Weather Service [Storm Data Documentation][1]

- National Climatic Data Center Storm Events [FAQ][2]

[1]: https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf
[2]: https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf

This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

In this analysis, I compared the impact on public health (number of injury and mortality) and amount of economic damage (damage on property and crop).

In the aspect of pulblic health, "tornado" has by far the greatest impact in both number of injuries and number of fatalities.
On the other hand, "flood" is the biggest cause of economic damage, especially in damage on property. In addition, "Drought" is the leading cause of ecomonic crop damage.


# Data Processing

First, download the data on your working directory.
Next, read the storm data to a R. This process may take some second. 
Please be patient!

```{r readSotrmFile,cache=TRUE}
if(!file.exists("stormData.csv.bz2")) {
        fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
        download.file(url = fileUrl, destfile = "stormData.csv.bz2", method = "curl")
}
stormdata <- read.csv(file = "stormData.csv.bz2")
```

There are `r ncol(stormdata)` fields in this file.

```{r colNames,cache=TRUE}
names(stormdata)
```

The data from 2000 to present was extracted because there are too many 
variations in `EVTYPE` (the name of the types) field. 

- [Database datail](http://www.ncdc.noaa.gov/stormevents/details.jsp?type=collection)

```{r extract,  results='hide',cache=TRUE}
library(lubridate)
stormdata$BGN_DATE <- mdy_hms(stormdata$BGN_DATE) # Convert to POSIXct format
stormdata <- stormdata[stormdata$BGN_DATE >= "2000-01-01",] # Subsetting
```


# Q1) Which type of events are most harmful with respent to population health?

## Variables

The following fields are related in this analysis.

- `EVNAME`: name of the event

- `FATALITIES`: the number of the fatalities in its recode.

- `INJURIES`: the number of the injuries in its recode.


First, the sum of the fatalities and injuries in each event name was calculated.

## Reshape the data

```{r reshapeHealth, cache=TRUE, results='hide'}
library(plyr)
healthsum <- ddply(stormdata, .(EVTYPE), summarise,
                fatalities = sum(FATALITIES),
                injuries = sum(INJURIES))
```

These data frame are reorderd by the number of fatalities and injuries,
and the top 20 and the sum of other cases are extracted.

The order of factor are arranged by the number of injuries and fatalities.

```{r reorderHealth,cache=TRUE}
library(plyr)
injOrd <- arrange(healthsum, desc(injuries))[1:20,c(1,3)]
injOrd$EVTYPE <- factor(injOrd$EVTYPE,
                        levels = injOrd[order(injOrd$injuries),"EVTYPE"])

fataOrd <- arrange(healthsum, desc(fatalities))[1:20, c(1,2)]
fataOrd$EVTYPE <- factor(fataOrd$EVTYPE,
                        levels = fataOrd[order(fataOrd$fatalities),"EVTYPE"])
```

## Result (1)
The number of injuries and fatalities from 2000 through 2011.

```{r plotting, fig.width=10,cache=TRUE}

library(ggplot2)
injOrd$injuries <- as.numeric(injOrd$injuries)
fataOrd$fatalities <- as.numeric(fataOrd$fatalities)

g1 <- ggplot(injOrd, aes(x = EVTYPE, y = injuries)) +
        geom_bar(stat = "identity") +
        ggtitle("INJURIES")+
        ylab(NULL) +
        xlab(NULL) +
        coord_flip()

g2 <- ggplot(fataOrd, aes(x = EVTYPE, y = fatalities)) +
        geom_bar(stat = "identity") +
        ggtitle("FATALITIES")+
        ylab(NULL) + 
        xlab(NULL) +
        coord_flip()

library(grid)
grid.newpage()
pushViewport(viewport(layout=grid.layout(1, 2)))
print(g1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(g2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

Tornadoes outnumbers the other cases in both injuries and fatalities. 
The most harmful event seems to be a tornadoes.
Although the number of injuries because of ECESSIVE HEAT is relatively
small comparing to that of TORNADO, it may ealily lead to death.


*******************************************************************************


# Q2) Which type of events have the greatest economic consequence?

## Variables

I'm going to compare the amount of "crop damage" and "property damage" among
each types of events.
Following culumns are associated in this analysis.

- `EVTYPE` : event type

- `CROPDMG` : crop damage

- `CROPDMGEXP` : crop damage exponent

- `PROPDMG` : property damage

- `PROPDMGEXP` : property damage exponent

The coloumn `CROPDMGEXP` and `PROPDMGEXP` contain several factors, but most of
them are not used. Then `K`, `M`, `B` seems to mean "thousand", "million",
"billion".

```{r summaryCrpAndProp}
summary(stormdata$CROPDMGEXP)
summary(stormdata$PROPDMGEXP)
```

For example, a record whose `CRPDMG` field has `12` and `CRPDMGEXP` field has
`M` says that the amount of crop damage is 12,000,000. 

In order to change this notation into real number,
`K`, `M` and `B` in exponent fields are replaced by character
`1`, `1000`, `1000000`, then `CRPDMG` and `PROPDMG` fields are replaced
by absolute amount of damage. (unit: million) 

Although a number of records has lack blank in `CROPDMGEXP` and `PROPDMGEXP`
field, these records are ingnored as `NA`s in this analysis.

## Reshape the data

```{r replace, cache=TRUE}
stormdata$CROPDMGEXP <- gsub("K", "0.001", stormdata$CROPDMGEXP)
stormdata$CROPDMGEXP <- gsub("M", "1", stormdata$CROPDMGEXP)
stormdata$CROPDMGEXP <- gsub("B", "1000", stormdata$CROPDMGEXP)

stormdata$PROPDMGEXP <- gsub("K", "0.001", stormdata$PROPDMGEXP)
stormdata$PROPDMGEXP <- gsub("M", "1", stormdata$PROPDMGEXP)
stormdata$PROPDMGEXP <- gsub("B", "1000", stormdata$PROPDMGEXP)

stormdata$CROPDMGEXP <- as.numeric(stormdata$CROPDMGEXP)
stormdata$PROPDMGEXP <- as.numeric(stormdata$PROPDMGEXP)

stormdata$CROPDMG <- stormdata$CROPDMG * stormdata$CROPDMGEXP
stormdata$PROPDMG <- stormdata$PROPDMG * stormdata$PROPDMGEXP
```

The amount of total damage in property and crop in each event type are
calculated.

```{r reshapeDMG,cache=TRUE}
library(plyr)
damage <- ddply(stormdata, .(EVTYPE), summarise,
                crop = sum(CROPDMG, na.rm = TRUE),
                prop = sum(PROPDMG, na.rm = TRUE))
```

Total amount of damage was ...

```{r reorderDMG, cache=TRUE}
total <- damage$crop + damage$prop
damage <- cbind(damage, total)

damage <- arrange(damage, desc(total))[1:20,]

damage$EVTYPE <- factor(damage$EVTYPE,
                        levels = damage[order(damage$total),"EVTYPE"])
```

## Result (2)

```{r plottingDMG, fig.width=10}
library(reshape2)
meltdamage <- melt(damage, id="EVTYPE", measure.vars = c("crop", "prop"))
meltdamage$value <- meltdamage$value/1000

library(ggplot2)
ggplot(meltdamage, aes(x = EVTYPE, y = value, fill = variable)) +
        geom_bar(stat = "identity") +
        ggtitle("Damage") +
        ylab("unit: billions") +
        xlab(NULL) +
        coord_flip() 
        
```

Flood has the greatest impact on ecomonic damage, followed by Hurrycane/typhoon and Storm surge. In many cases, the amount of damage on property are larger than that on crop, then the total amount of damage tend to reflect mainly the amount of property damage. In the aspect of crop damage, Drought is the leading cause.